---
title: "hw_3"
author: "Ray Hunter"
date: "2023-05-10"
output:
  html_document:
    code_folding: hide
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(janitor)
library(here)
library(pracma)



```

##  {.tabset}

### 1

*The data set HW3 data.csv provides price (in \$) and quantity (in
number of EVs) estimates of demand per year for both DAC and non-DAC
groups. Run linear regressions (with an intercept) to estimate the
demand curves for DAC and non-DAC consumers. What are 2-3 reasons you
can think that may explain the differences in demand you estimate?*

```{r}
#reading in the data and cleaning names 

hw3_data <- read.csv(here("HW3_data.csv")) %>% 
  clean_names() %>%  
  rename(price = price_dollars)  %>% 
  group_by(price) %>% 
  #adding a new colomn for aggregate quantity 
  mutate(ag_dem=sum(q_dac + q_non_dac)) %>% 
  pivot_longer(
    cols= c(q_dac:ag_dem),
    names_to = "type",
    values_to = "quantity")
  





#creating an lm of dac
dac_lm <- lm(price ~ quantity, data = filter(hw3_data, type =="q_dac"))


#lm of non dac
non_dac_lm <- lm(price ~ quantity, data = filter(hw3_data, type =="q_non_dac"))


```




```{r}
#list of functions made



###################################################################


#calculate Quantity demanded from intercept, slope, and price 
dem_q <- function(int, P, slope){
  Q = (int -P) / - slope
  return(Q)
}
###################################################################

#calculate price of quantity from intercept, slope, and price 
dem_p <- function(int, slope, Q){
  P = int + slope*Q
  return(P)
}


###################################################################



#new function solving quantity supplied at a given price 
supp_q <- function(int2, P, slope2){
  Q = (int2 - P)/slope2
    return(Q)
} 


###################################################################
#calculate price of quantity from intercept, slope, and price 
supp_p <- function(int2, slope2, Q){
  P = int2 + slope2*Q
  return(P)
}

###################################################################



#third function finding the difference between two other functions
F3 <- function(Q, int, slope, int2, slope2){
  z<- dem_p(int, slope, Q) - supp_p(int2, slope2, Q)
  return(z)
  }








```



need to add a kink: agg-d = ifelse(agg-d <0 non-dac, nonDAcdem, agg-d)

### need to add the two curves together here, not the data points 




Non-DAC's are willing to pay more for electric vehicles to keep their
communities clean than DAC's are. This is likely a result of Non-DAC's
tending to be composed of wealthier predominately white individuals that
have more disposable income. Therefore, they are willing ot pay more to
have a clean neighborhood than a DAC community member who makes less
money.

**DAC Demand**

$$ P=116,855-6.61Q $$

**Non-DAC Demand**

$$ P=158,034-2.73Q $$

### 2

*One kilogram (kg) of lithium is estimated to cause \$300 dollars worth
of damage in health costs to local communities living near mines. Each
EV requires a lithium ion battery containing 63kg of lithium. Assuming
there are no other externalities caused by EV production, what is the
marginal externality cost per EV produced?*

```{r}
mec_ev <- 300*63
mec_ev
```

the answer is `r mec_ev`

### 3

```{r}

#making a new df creating the aggregate demand 
ag_data <- data.frame(price = seq(0,160000, length.out=101)) %>% 
  
  #new colomn for  finding the quantity demanded at each price 
  mutate(q_dac = pmax( 0, dem_q(int= dac_lm$coefficients[[1]], P = price, slope=dac_lm$coefficients[[2]] ))) %>%
  mutate(q_non_dac = pmax(0, dem_q(int= non_dac_lm$coefficients[[1]], P = price, slope=non_dac_lm$coefficients[[2]] ))) %>%
  #adding up each of the quantities demanded at each price 
  mutate(q_agg = q_dac + q_non_dac) %>% 
  pivot_longer(
    cols= c(q_dac:q_agg),
    names_to = "type",
    values_to = "quantity")

 # ag_data[ag_data == 0] <- NA

######################  Agg dem equation  ###############################################

ag_lm <- lm(price ~ quantity, data = filter(ag_data, type =="q_agg"))

ag_lm_lower <- lm(price ~ quantity, data = filter(ag_data, type =="q_agg", price <= dac_lm$coefficients[[1]]))
ag_lm_higher <- lm(price ~ quantity, data = filter(ag_data, type =="q_agg", price >= dac_lm$coefficients[[1]]))

```

```{r, fig.cap = "This graph here shows the three different demand curves for EV's in DAC's, non-DAC's, and their aggregate demand. It shows that DAc communities have a lower demand and non-DAC's have a higher demand. The kink in the aggregate demand represents the aggregate demand mirroring the non-DAC demand at. agiven price until the DAC demand also starts."}

#creating a ggplot of both the lm
ggplot(ag_data, aes(x= quantity, y = price, color = type)) +
  # geom_smooth(se=FALSE) +
  labs(x = "Quantity of EVs", y = "Price ($)") +
geom_line() +

  # stat_smooth(method = "lm", se=FALSE) +
# geom_hline(yintercept=50000, color = "black") +
  theme_minimal()


```

*What is the aggregate annual demand curve for EVs?*

$$ P = 151,383 -2.036 Q $$

*What is the supply curve for EVs?*

```{r}
#quantity at 50,000
q_market <- dem_q(ag_lm$coefficients[1], 50000, ag_lm$coefficients[2])


#slope of the supply 
sup_slope <- (0-50000)/(0 - q_market)



```


**Supply Curve:**
$$ P =   1.004107 Q $$

*What is the "benefit" to consumers under the status quo?*

```{r}



#calculating the consumer benefit triangle 
cons_ben <- 0.5*(ag_lm$coefficients[1] - 50000)*q_market



#calculating producer benefit 
prod_ben <- 0.5*(50000-0)*q_market



#calculating the environmental cost
market_envtl_cost <- q_market* mec_ev


#calculating the consumer benefit triangle of the dac community manually 
dac_ben <- 0.5*(dac_lm$coefficients[1] - 50000)* dem_q(dac_lm$coefficients[1], 50000, dac_lm$coefficients[2])
dac_ben

#take the aggregate benfit and subtract the dac benefit to find the non-dac benefit 

non_dac_ben <- 0.5*(non_dac_lm$coefficients[1] - 50000)* dem_q(non_dac_lm$coefficients[1], 50000, non_dac_lm$coefficients[2])
non_dac_ben


#try to do the integral instead when you have time 


```

The benefit to consumers is **$2,524,210,906**.

*What is the "benefit" to producers under the status quo?* 

The benefit to producer is **$ 1,244,887,764 **.

*What is the environmental cost under the status quo?*



The environmental cost is **$941,135,149  ** at the market equilibrium.



### 4

*How is the current consumer benefit divided between DAC and non-DACconsumers?*

The DAC benefit is  **$338,017,962** and the Non-DAC benefit is **$2,136,894,908** of the total consumer benefit.



### 5

*Derive the optimal EV tax (in \$ per vehicle) to internalize the lithium mining externality. Noting that recent research has shown that DACs are far more likely to contain mining activity, assume that the mining externality is borne entirely by the DAC group. What would be the effects of this tax on:*

The optimal EV tax to internalize the lithium mining externality should be **$18,900** per vehicle 

**(a)** The amount of EVs produced and consumed 



```{r}
## The amount of EVs produced and consumed 
msc_eq_q <- fzero(function(Q) F3(Q, int= ag_lm$coefficients[1] , slope = ag_lm$coefficients[2], int2 = 18900 , slope2  = sup_slope), c(-10000000000, 1000000000000))$x
msc_eq_q



```



**(b)** The price of EVs

```{r}
#The price of EVs. plug in the demanded quantity to find price 
msc_eq_p <- dem_p(int= ag_lm$coefficients[1], Q = msc_eq_q, slope = ag_lm$coefficients[2])
msc_eq_p


```




**(c)** Overall welfare of non-DAC consumers 


```{r}
#finding the area of the triangle for non-DAC consumers 
#.5 * price * quantity demanded at MSC and aggregate demand equilibrium for the non-dac
non_dac_welfare <- 0.5*(non_dac_lm$coefficients[1] - msc_eq_p)*dem_q(non_dac_lm$coefficients[1], P = msc_eq_p, slope = non_dac_lm$coefficients[2])

non_dac_welfare

```




**(d)** Overall welfare of DAC consumers 


```{r}
##finding the area of the triangle for DAC consumers 
#.5 * price * quantity demanded at MSC and aggregate demand equilibrium for the dac
dac_welfare <- 0.5*(dac_lm$coefficients[1] - msc_eq_p)*dem_q(dac_lm$coefficients[1], P = msc_eq_p, slope = dac_lm$coefficients[2])

dac_welfare

```



**(e)** EV producers 

```{r}

#.5 x the equilibrium price * equilibrium quantity 
prod_welfare <- 0.5*(msc_eq_p - mec_ev)*msc_eq_q
prod_welfare


```




**(f)** Total environmental damage

```{r}
#environmental price multiplied by the quantity 
envt_damage <- mec_ev*msc_eq_q
envt_damage

```


**(g)** Total tax revenue generated


```{r}
#environmental price multiplied by the quantity 
tax_rev <- mec_ev*msc_eq_q
tax_rev

```





### 6

*Now, assume that all revenue from the EV tax will be redistributed to
the consumers in proportion to their pre-tax consumption. For example,
if 80% of the EVs were consumed by non-DAC consumers, then they get 80%
of the tax revenue. Additionally, consider that emerging scientific
evidence suggests the true environmental damages from lithium mining may
be much higher than \$300. For a range of values of external costs per
kg of lithium (\$350, \$400, \$450, and \$500 per kg), calculate the
effects of an EV tax on:*

**(a)** Overall welfare of non-DAC consumers

**(b)** Overall welfare of DAC consumers

**(c)** EV producers




### 7

*Now, consider the fact that the purchase of EVs not only leads to
lithium mining, but also helps mitigate climate change by replacing
gasoline cars. Suppose that climate change damages fall mostly outside
California, to a lesser extent on DAC consumers, and not at all on
non-DAC consumers. Qualitatively answer the following (in 1-3 sentences
each):*

**(a)** How does the presence of climate change damages from gasoline
cars affect the optimal EV tax in California?

**(b)** Assuming tax revenue does not get returned to consumers, are
non-DAC consumers more or less likely to support this new tax, relative
to the tax that only accounts for the mining externality? Why?

**(c)** Assuming tax revenue does not get returned to consumers, are DAC
consumers more or less likely to support this new tax, relative to the
tax that only accounts for the mining externality? Why?
